---
title: "PA1_template"
author: "coursera"
date: "12/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Reproducible Research: Peer Assigment 1

##Loading data

Frist we will unzip, load and view the data structure

```{r}
unzip("./activity.zip")
data <- read.csv("./activity.csv")
summary(data)
```

```{r}
head(data)
```

## Histogram of Total Number of steps taken each day

we will aggregate the steps(value) for each corresponding dates.

```{r}
steps_per_day <- aggregate(steps ~ date, data, sum, na.rm= TRUE)
hist(steps_per_day$steps, col = "orange", xlab = "Steps", ylab = "Frequency", main = "Steps taken each date")
```

##Mean and Median of steps taken each date

we will now compute the mean and median

```{r}
steps_mean <- mean(steps_per_day$steps)
steps_median <- median(steps_per_day$steps)

cat(" Mean is", steps_mean , 
    "Median is ", steps_median)
```


## Time series plot

Time series plot of the average number of steps taken

```{r}
avg_steps_date <- aggregate(steps ~ date, data, mean, na.rm=TRUE)

plot(avg_steps_date$steps,  type = "l")

```

Maximum number of steps in 5 minutes interval

```{r}
steps_for_interval <- aggregate(steps ~ interval, data, mean, na.rm = TRUE)
plot(steps ~ interval , steps_for_interval, type ="l")
```

```{r}
Max_interval_with_steps <- steps_for_interval[which.max(steps_for_interval$steps),]
cat("Max average was at" , Max_interval_with_steps$interval, "th Interval")
```


## Working aroun dthe missing values

First lets compute the missing values in the dataset

```{r}
colSums(is.na(data))
```

There are 2304 miissing values in the steps column.
We will replace these missing values with average steps by interval.

```{r}
m <- mean(steps_for_interval$steps)
```

Now lets replace the value
```{r}
data1 <- data
missingIndex <- is.na(data[,1])

data1[missingIndex,1] <- m
head(data1)
```
steps taken each day atter computing the missing value

```{r}
steps_nona <- aggregate(steps ~ date, data1, sum )

hist(steps_nona$steps, col = "blue", xlab = "Steps", ylab = "Days", main = "Number of steps taken after replacing missing values")
```

## Difference in Steps by Interval between weekday and weekend

Conver the date column into date class and inmport dplyr library

```{r}
data1$date <- as.Date(data1$date)
library(dplyr)
```


Now lets add a column to the dataset to determine whethere the date is weekday or weekend


```{r}
data2 <- data1 %>%
  mutate(dateType = ifelse(weekdays(data1$date) == "Saturday" | weekdays(data1$date) == "Sunday", "Weekend", "Weekday"))
head(data2)
```

Now lets sums the steps for Interval and dateType

```{r}
dataSum <- data2 %>%
  group_by(dateType, interval)%>%
  summarize(avg = sum(steps))

head(dataSum)

```

Plot the graphs

```{r}
library(lattice)

xyplot(avg ~ interval |dateType, dataSum, type ="l", layout = c(1,2))
```











